---
title: "Recession-Forecast"
author: "Jai Vrat Singh"
date: "23/09/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r warning=FALSE, message=FALSE}
library(xts)
library(PerformanceAnalytics)
library(tseries)
library(forecast)
library(knitr)
library(kableExtra)
library(ggplot2)
```

Set working directory

```{r}
setwd("/Users/jvsingh/work/github/R-codes/recession")

direct.df <- read.csv(file = "./data/2019-08.csv", header = TRUE, stringsAsFactors = FALSE, check.names = FALSE)
direct.df <- direct.df[-1, ] #The first row is transform etc..

#Date is in month/date/year format
direct.df$sasdate <- as.Date(direct.df$sasdate, "%m/%d/%Y")

#get into time series as it will be earier to work with
direct.df.xts <- xts(direct.df[,-1], direct.df$sasdate)

#Replace missing values by the latest observation. Please note that there may be missing values in "real" decision scenarios which may not be 
#caught by one time experiment. ie in case if we are doing real backtests
#direct.df.xts <- zoo::na.locf(direct.df.xts)
```


```{r}
plot.xts(direct.df.xts, legend.loc = "bottomright")
```


As seen from the plot many series seem to ne non-stationary. We need to check for seasonality as well.

```{r}
#install.packages("visdat")
library(visdat)
vis_dat(data.frame(direct.df.xts))
```

```{r}
 apply(direct.df.xts, 1, function(elem) sum(is.na(elem)))
```
Seems that there are many missing till 1959-12-01. Let us drop them

```{r}
direct.df.xts <- window(direct.df.xts,start = as.Date("1959-12-01") + 1)
vis_dat(data.frame(direct.df.xts))
```

There are 5 columns which have got NA's. 


            RPI         W875RX1 DPCERA3M086SBEA         RETAILx          INDPRO         IPFPNSS         IPFINAL         IPCONGD 
              0               0               0               0               0               0               0               0 
       IPDCONGD        IPNCONGD         IPBUSEQ           IPMAT          IPDMAT          IPNMAT       IPMANSICS       IPB51222S 
              0               0               0               0               0               0               0               0 
        IPFUELS          CUMFNS         CLF16OV          CE16OV          UNRATE        UEMPMEAN         UEMPLT5       UEMP5TO14 
              0               0               0               0               0               0               0               0 
       UEMP15OV       UEMP15T26        UEMP27OV         CLAIMSx          PAYEMS          USGOOD   CES1021000001          USCONS 
              0               0               0               0               0               0               0               0 
         MANEMP         DMANEMP        NDMANEMP          SRVPRD           USTPU        USWTRADE         USTRADE          USFIRE 
              0               0               0               0               0               0               0               0 
         USGOVT   CES0600000007         AWOTMAN          AWHMAN           HOUST         HOUSTNE         HOUSTMW          HOUSTS 
              0               0               0               0               0               0               0               0 
         HOUSTW          PERMIT        PERMITNE        PERMITMW         PERMITS         PERMITW         AMDMNOx         AMDMUOx 
              0               0               0               0               0               0               0               0 
           M1SL            M2SL          M2REAL           AMBSL        TOTRESNS       NONBORRES        BUSLOANS          REALLN 
              0               0               0               0               0               0               0               0 
        S&P 500     S&P: indust        FEDFUNDS           CP3Mx           TB3MS           TB6MS             GS1             GS5 
              0               0               0               0               0               0               0               0 
           GS10             AAA             BAA       COMPAPFFx        TB3SMFFM        TB6SMFFM          T1YFFM          T5YFFM 
              0               0               0               0               0               0               0               0 
        T10YFFM          AAAFFM          BAAFFM         EXSZUSx         EXJPUSx         EXUSUKx         EXCAUSx      WPSFD49207 
              0               0               0               0               0               0               0               0 
     WPSFD49502         WPSID61         WPSID62       OILPRICEx          PPICMM        CPIAUCSL        CPIAPPSL        CPITRNSL 
              0               0               0               0               0               0               0               0 
       CPIMEDSL     CUSR0000SAC     CUSR0000SAD     CUSR0000SAS        CPIULFSL   CUSR0000SA0L2   CUSR0000SA0L5           PCEPI 
              0               0               0               0               0               0               0               0 
DDURRG3M086SBEA DNDGRG3M086SBEA DSERRG3M086SBEA   CES0600000008   CES2000000008   CES3000000008           MZMSL          INVEST 
              0               0               0               0               0               0               0               0 
      CMRMTSPLx             HWI       HWIURATIO         BUSINVx        ISRATIOx        NONREVSL          CONSPI   S&P div yield 
              1               1               1               1               1               1               1               1 
    DTCOLNVHFNM        DTCTHFNM    S&P PE ratio         VXOCLSx         ANDENOx        UMCSENTx        TWEXMMTH          ACOGNO 
              1               1               3              30              97             144             156             386 


ACOGNO : Value of Manufacturers' New Orders for Consumer Goods Industries (386 NA) Starts from  1992
TWEXMMTH: Trade Weighted U.S. Dollar Index: Major Currencies, Goods       (156) Starts from  1973
UMCSENTx: University of Michigan: Consumer Sentiment                      (144) Starts from Feb 1978







