---
title: "ICA"
author: "Jai Vrat Singh"
date: "06/07/2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

** Independent Component Analysis **

* Assume that two individual components are mixed by the followgin matrix


$\begin{bmatrix}
   A_0 
\end{bmatrix}
=
\begin{bmatrix}
    2 & 3 \\
    2 & 1 
\end{bmatrix}$


that is

$\begin{bmatrix}
    x_1 \\
    x_2 
\end{bmatrix}
=
\begin{bmatrix}
    2 & 3 \\
    2 & 1 
\end{bmatrix}\begin{bmatrix}
    s_1 \\
    s_2 
\end{bmatrix}$



```{r cars}
set.seed(144)
N = 10000
dat <- data.frame(s1 = runif(N) - 0.5, s2 = runif(N) - 0.5)
head(dat)
```

## Including Plots

You can also embed plots, for example:

```{r, echo=FALSE}
library(ggplot2)
ggplot(dat, aes(x=s1, y=s2)) +
    geom_point(size=0.5)      # Use hollow circles
```



```{r}
#Mixture or transformation matrix
A <- rbind(c(2,3),
           c(2,1))
dat.m <- as.matrix(dat)
dim(dat.m)

head(dat.m)

# X = (x1, x2)
X = A %*% t(dat.m)
dat.tr = as.data.frame(t(X))
colnames(dat.tr) <- c("x1", "x2")

ggplot(dat.tr, aes(x=x1, y=x2)) +
    geom_point(size=0.5) +     # Use hollow circles
    geom_abline(intercept = 1, slope = A[1,1]/A[2,1], color="red") +    #I calculated this 1
    geom_abline(intercept = -2/3, slope = A[2,2]/A[1,2], color="green") #I Calculated this -2/3

```

Now the variables are not independent. Also note that the slopes of two edge lines are nothing but column vectors of the A matrix

