---
title: "Arch-garch-modelling"
author: "Jai Vrat Singh"
date: "11/09/2018"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(fGarch)
```

## R Markdown


```{r cars}
ret.what  <- list(date=numeric(), returns=numeric())
ret.widths <- c(12, 8)
strip.white <- c(TRUE, TRUE)

datF = scan(file = "http://faculty.chicagobooth.edu/ruey.tsay/teaching/fts2/m-3m4603.txt",
             what = ret.what, strip.white = strip.white)
datF = as.data.frame(datF)


#Above are simple returns, Convert returns to logReturns
datF[["logRet"]] = log(1- datF$returns)

head(datF)
```

## Evidence of ARCH effect


```{r}
plot(datF$logRet, type = "o", pch = ".")
```

This plot reflects a pattern around the line logRet = 0, We need to look into ACF and PACF to get lagged correlations and its exact nature before the ARCH effect can be detected.


```{r}
par(mfrow=c(2,1))
acf(datF$logRet)
acf(datF$logRet, type = "partial")
```


ACF till lags 12 => MA(12)
PACF till 3, 12 => AR(3), AR(12)

Try these

```{r}
#AR(12)
logret.fit = arima(x = datF$logRet, order = c(12,0,0))
logret.fit
#aic = -1779.14

#AR(3)
logret.fit = arima(x = datF$logRet, order = c(3,0,0))
logret.fit
#aic = -1781.7

#MA(12)
logret.fit = arima(x = datF$logRet, order = c(0,0,12))
logret.fit
#aic = -1780.48

```

Lowest AIC is for AR(3) => seems it is best fit

## Residue analysis of above fit

```{r}
#logret.fit$residuals
qqnormPlot(logret.fit$residuals)
```
```{r}
acf(logret.fit$residuals, lag.max = 20)

```

```{r}
acf(logret.fit$residuals, type = "partial", lag.max = 20)
```



Trying various models and Residual analysis, seems that MA(12) is the best fit model.

```{r}
lags <- c(1:20)
sapply(lags, FUN = function(lag) {
                    test <- Box.test(logret.fit$residuals, lag = lag, type = c("Box-Pierce", "Ljung-Box"), fitdf = 0)
                    test$p.value
})
```

P-values are very high => the residual is white noise.

## ARCH effect on residuals

```{r}
par(mfrow=c(2,2))
acf(logret.fit$residuals, lag.max = 20)
pacf(logret.fit$residuals, lag.max = 20)
acf(logret.fit$residuals^2, lag.max = 20)
pacf(logret.fit$residuals^2, lag.max = 20)
```

ACF and PACF of residuals clearly show conditional heteroscedasticity

Let us do LB test on residuals^2

```{r}
lags <- c(1:20)
sapply(lags, FUN = function(lag) {
                    test <- Box.test(logret.fit$residuals^2, lag = lag, type = c("Box-Pierce", "Ljung-Box"), fitdf = 0)
                    test$p.value
})
```

P-values are less than 0.05 => means that shocks are not independent. There is ARCH effect in almost all the lags.


## PACF of squared log returns

```{r}
pacf(datF$logRet^2, lag.max = 20)
```

Seems and ARCH(2) model can be fitted

$r_t = \mu + a_t$

$a_t = \sigma_t \epsilon_t$

$\sigma_t^2 = \alpha_0 + \alpha_1 a_{t-1}^2 + \alpha_2 a_{t-2}^2$

Fit the Model

```{r}
library(fGarch)
garch2.fit <- garchFit(~garch(2,0), data = datF$logRet, trace = FALSE)
garch2.fit
```

$\mu = -0.0169129$

$\alpha_0 = \omega = 0.0035752$

$\alpha_1 = 0.1125097$

$\alpha_2 = 0.0985089$

So the Model is:

$r_t = -0.0169129 + a_t$

$\sigma_t^2 =  0.0035752 + 0.1125097 a_{t-1}^2 + 0.0985089 a_{t-2}^2$

$a_t = \sigma_t \epsilon_t$

```{r}
resid <- residuals(garch2.fit)
par(mfrow=c(2,1))
acf(resid, lag.max = 20)
pacf(resid, lag.max = 20)
```
```{r}
qqnormPlot(resid)
```


